---
title: "Chapter 2: Building Blocks Neural Networks"
output: html_notebook
---

# Example Classifier

Build a classifier for the MNIST dataset: the `"hello world"` of deep learning.

```{r}
library(keras)

mnist <- dataset_mnist()
train_images <- mnist$train$x
train_labels <- mnist$train$y
test_images <- mnist$test$x
test_labels <- mnist$test$y
```

```{r}
str(train_images)
str(train_labels)
str(test_images)
str(test_labels)
```



Now set up the architecture of the neural network.

```{r}
network <- keras_model_sequential() %>% 
    layer_dense(units = 512, 
                activation = "relu", 
                input_shape = c(28 * 28)) %>% 
    layer_dense(units = 10, 
                activation = "softmax")
```

Now need to add a loss function, an optimiser, and metrics to monitor during training and testing.

```{r}
network %>% 
    compile(
        optimizer = "rmsprop", 
        loss = "categorical_crossentropy", 
        metrics = c("accuracy")
    )
```

Now reshape and scale the data.

```{r}
train_images <- array_reshape(train_images, 
                              c(60000, 28 * 28))
train_images <- train_images / 255

test_images <- array_reshape(test_images, 
                             c(10000, 28 * 28))
test_images <- test_images/255
```

And change labels to required structure.

```{r}
train_labels <- to_categorical(train_labels)
test_labels <- to_categorical(test_labels)
```

Now train the network using `keras::fit()`.

```{r}
network %>% 
    fit(train_images, 
        train_labels, 
        epochs = 5, 
        batch_size = 128)
```

```{r}
metrics <- network %>% 
    evaluate(test_images, 
             test_labels)
metrics
```

```{r}
network %>% 
    predict_classes(test_images[1:10, ])
```

#  Data Structures

```{r}
# 0D tensor
# R doesn't exactly have an equivalent of a scalar/0D tensor, but a vector
# of length 1 is about the same
zeroD <- 1

# 1D tensor
# just a vector
oneD <- c(1, 4, 9, 16)

# 2D tensor
# this is a matrix
twoD <- matrix(rep(0, 3 * 5), nrow = 3, ncol = 5)
dim(twoD)

# 3D and above tensors
# pack matrices into an array, form a cube, which is a 3D tensor
threeD <- array(rep(0, 2 * 3 * 2), dim = c(2, 3, 2))
str(threeD)
dim(threeD)
# can then put 3D tensors in an array to make 4D tensor, those to make 5D etc.
```

```{r}
train_images <- mnist$train$x
train_labels <- mnist$train$y
test_images <- mnist$test$x
test_labels <- mnist$test$y
```

An example of one image.

```{r}
plot(as.raster(train_images[5, , ], max = 255))
```

More examples of tensor slicing.

```{r}

```


