---
title: "Chapter 3: Getting Started with Neural Networks"
output: html_notebook
---

# Classifying Movie Reviews

```{r}
library(keras)
```

```{r}
imdb <- dataset_imdb(num_words = 10000)
```

```{r}
# this uses the %<-% multiple assignment operator from the zeallot package to 
# take the elements of a list and assign them to several objects at once
c(c(train_data, train_labels), c(test_data, test_labels)) %<-% imdb
```

Can use `dataset_imdb_word_index()` to translate these back into English.

```{r}
word_index <- dataset_imdb_word_index()
reverse_word_index <- names(word_index)
names(reverse_word_index) <- word_index
decoded_review <- sapply(train_data[[1]], function(index) {
    word <- if (index >= 3) reverse_word_index[[as.character(index - 3)]]
    if (!is.null(word)) word else "?"
})
decoded_review
```

Now one-hot encode them ready for the network.

```{r}
vectorise_sequences <- function(sequences, dimension = 10000L) {
    results <- matrix(0, nrow = length(sequences), ncol = dimension)
    for (i in seq_along(sequences)) {
        results[i, sequences[[i]]] <- 1
    }
    results
}

x_train <- vectorise_sequences(train_data)
x_test <- vectorise_sequences(test_data)
```

Change labels to numeric type.

```{r}
y_train <- as.numeric(train_labels)
y_test <- as.numeric(test_labels)
```

Create network structure and compile with loss function and optimiser.
```{r}
model <- keras_model_sequential() %>%
    layer_dense(units = 16, activation = "relu", input_shape = c(10000)) %>%
    layer_dense(units = 16, activation = "relu") %>%
    layer_dense(units = 1, activation = "sigmoid")

model %>% compile(
    optimizer = "rmsprop",
    loss = "binary_crossentropy",
    metrics = c("accuracy"))
```

Set aside a portion of the data for validation.

```{r}
val_indices <- 1:10000
x_val <- x_train[val_indices,]
partial_x_train <- x_train[-val_indices,]
y_val <- y_train[val_indices]
partial_y_train <- y_train[-val_indices]        
```

Fit the model.

```{r}
history <- model %>% fit(
    partial_x_train,
    partial_y_train,
    epochs = 20,
    batch_size = 512,
    validation_data = list(x_val, y_val))
```

```{r}
str(history)
plot(history)
```

Model overfits quite badly. Try a different setup to improve performance.

```{r}
model <- keras_model_sequential() %>%
    layer_dense(units = 16, activation = "relu", input_shape = c(10000)) %>%
    layer_dense(units = 16, activation = "relu") %>%
    layer_dense(units = 1, activation = "sigmoid")
model %>% compile(
    optimizer = "rmsprop",
    loss = "binary_crossentropy",
    metrics = c("accuracy")
)
model %>% fit(x_train, y_train, epochs = 4, batch_size = 512)
results <- model %>% 
    evaluate(x_test, y_test)
```

```{r}
results
```

Just reducing the number of epochs gets much better performance. Now use that model to predict on test data.

```{r}
model %>% 
    predict(x_test[1:10, ])
```

Varying confidence on these examples.

